Practical Performance Evaluation of Entity Resolution Algorithms: Lessons Learned at PatentsView.org

### 2022-07-20 paper/main.tex

@@

\author[1,2]{Olivier Binette}
\author[1]{Youngsoo Baek}
\author[2]{Aida York}
\author[2]{Emma Hickerson}
\author[2]{Sarvo Madhavan}
\author[2]{Christina Jones}
\affil[1]{Department of Statistical Science, Duke University}
\affil[2]{American Institutes for Research}

@@

\begin{abstract}
This paper introduces a novel evaluation methodology for entity resolution algorithms. It is motivated by PatentsView.org, a publicly-available patent data analysis tool which disambiguates patent inventors based on an entity resolution algorithm. We provide a data collection methodology and performance estimators which account for sampling biases. Our approach is simple, practical and principled -- key characteristics which allow us to paint the first representative picture of PatentsView's disambiguation performance. This is used to inform PatentsView's users of the reliability of the data and to allow the comparison of competing disambiguation algorithms.
\end{abstract}

\section{Introduction}
Entity resolution is the task of identifying records in a database which refer to the same entity. An entity may be a person, a company, an object or an event. Records are assumed to contain partially identifying information about these entities. When there is no unique identifier (such as a social security number) available for all records, entity resolution becomes a complex problem which requires sophisticated algorithmic solutions \citep{Herzog2007, Christen2012, Dong2015, Ilyas2019, Christophides2019, Christen2019, Papadakis2021, Binette2022a}.

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions to inventors, where a single person's name may be spelled in different ways on two applications, and where two different inventors with the same name may be difficult to distinguish. Much research has ben done within the fields of economics, computer science and statistics to disambiguate inventor mentions \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

@@

The entity resolution evaluation problem is to extrapolate from observed performance on small samples to real performance in a database with millions of records. \cite{Wang2022} refers to this as bridging the ``reality-ideality gap" in entity matching, where high performance on benchmark datasets often does not translate in the real world. Here, performance may be defined as any combination of commonly used evaluation metrics for entity resolution, such as precision and recall, cluster homogeneity and completeness, rand index, or generalized merge distance \citep{Maidasani2012}. These metrics can be computed on benchmark datasets for which we have a ground truth disambiguation. However, the key evaluation problem is to obtain estimates which are representative of performance on the full data for which no ground truth disambiguation is available. This is challenging for the following reasons.

First, entity resolution problems scale non-linearly. While it may be easy to disambiguate a small dataset, the opportunity for errors grows \textit{quadratically} in the number of records. As such, we may observe good performance of an algorithm on a small benchmark dataset, while the true performance of on the entire data may be something else entirely.\footnote{This particular effect of dataset size in entity resolution is explored in \cite{Draisbach2013} in the context of choosing similarity thresholds.} This is a problem that PatentsView.org currently faces. Despite encouraging performance evaluation metrics on benchmark datasets, with nearly perfect precision and recall reported in the latest methodological report \citep{Monath2021}, the data science team at AIR observes lower real-world accuracy.

A second problem is large class imbalance in entity resolution \citep{Marchant2017}. Among all pairs of records, only a small fraction (usually much less than a fraction of a percent) refer to the same entity. The vast majority of record pairs are not a match. This makes it difficult to evaluate performance through random sampling.

@@

Our approach, detailed in sections \ref{sec:approach} and \ref{sec:proposed_estimators}, addresses these challenges by putting forward cluster-based expressions for performance metrics which reflect various sampling schemes. Each of these representations immediately suggests simple estimators which properly account for the above issues.

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we have carried out a simple experiment which is described in detail in appendix \ref{appendix:example}. In short, we considered evaluating a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark data sets are obtained and used. For reference, it is known that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

@@

Pairwise precision and recall metrics were first reported in \cite{Newcombe1959}, with \cite{Bilenko2003, Christen2007} emphasizing the importance of precision-recall curves for algorithm evaluation. Now, there are issues with the use of pairwise precision and recall in entity resolution applications, such as the large relative importance of large clusters. As such, other clustering metrics have been proposed, including cluster precision and recall, cluster homogeneity and completeness, and the generalized merge distance \citep{Michelson2009, Menestrina2010, Maidasani2012, Barnes2015}. Important practical issues regarding the use of aggregate metrics are discussed in \cite{Hand2018}.

@@

@@

Our approach to estimating performance metrics is based on the use of benchmark datasets which are already existing or which can be collected in a cost-effective way. This data contains entity clusters corresponding to sampling records and recovering all associated instances, to directly sampling clusters, or to sampling blocks. For each of these sampling processes, we propose estimators which correct for the issues discussed in section \ref{sec:challenges}, which are nearly unbiased, and which are easy to use in practice.

@@

@@

The rest of the paper is organized as follows. In section \ref{sec:data_methods}, we introduce motivating benchmark datasets, we describe our data collection process for evaluation, we introduce the proposed estimators, and we summarize our simulation study used to validate the performance of our estimators. Section \ref{sec:results} then shows the results -- our performance estimates and results from the simulation study. Finally, section \ref{sec:discussion} summarizes the paper and discusses open research directions.

@@

In this section, we review motivating datasets used at PatentsView, we introduce our hand-disambiguation methodology, we describe our proposed performance metric estimators, and we describe our simulation framework used to compare estimators.

@@

\paragraph{Israeli Inventors Benchmark}
\cite{trajtenberg2008identification} disambiguated the US patents of Israeli inventors which were granted between 1963 and 1999. A total of 6,023 Israeli inventors were identified for this time period with 15,310 associated patents.

@@

Based on an original dataset from \cite{gu2008march}, \cite{Li2014} disambiguated the patent history (between 1975 and 2010) of 95 US inventors.

@@

In addition to the above benchmark datasets, we have carried out a data collection experiment which was motivated by the evaluation, using the proposed estimators discussed in section \ref{sec:proposed_estimators}, of the current PatentsView inventor disambiguation.

In total, 100 inventors were sampled with probability proportional to their number of granted patents. This was done by sampling inventor mentions uniformly at random and recovering all patents for a given inventor. These inventor mentions were from US patents granted between 1976 and December 31, 2021.

@@

Note that our data collection methodology is biased towards PantentsView's current disambiguation. In cases where no errors were found, the current disambiguation was assumed to be correct. Performance estimates based on this data might therefore be slightly optimistic.

@@

\begin{remark}
    Lemma \ref{lemma:record_sampling_representation} and lemma \ref{lemma:cluster_sampling_representation} can be generalized to apply to any performance metric which can be expressed as a sum $\sum_{c \in \mathcal{C}} h(c, \widehat{\mathcal{C}})$, for some function $h$, or as a function of such sums. For instance, the use of the so-called \textit{cluster} precision and \textit{cluster} recall \citep{}, or of cluster homogeneity and completeness \citep{}, can be more appropriate in the presence of large clusters. We leave these generalizations as extensions of our work.
\end{remark}

@@

All of the expressions for precision and recall in lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are either population means or ratios of population means. As such, they can be estimated using sample means and ratios of sample means. For readability, we present here a generic formula for an approximately unbiased estimator of the ratio of means, and then specify the needed quantities for each representation below. The estimator applies a first order bias correction to the ratio of sample means, based on Taylor approximation. Approximate confidence intervals can be computed based on the variance estimator of the ratio of sample means by Taylor approximation, assuming the corrected estimator has small bias \citep{fuller2011sampling}. The generic formula for estimating the ratio of $T$-sized ``population'' (of records/clusters/blocks) means of the form
\begin{align}
    E=\frac{1/T\sum_{i=1}^{T} B_i}{1/T\sum_{i=1}^T A_i},
\end{align}
assuming we have sampled $n$ elements (records/clusters/blocks), is
\begin{align}\label{eq:bias_adjustment}
    \widehat{E}=\frac{\bar{B}_n}{\bar{A}_n}\left\{1 + \frac{\theta_{n,T}}{n(n-1)}\frac{1}{\bar{A}_n}\sum_{s=1}^n A_s
    \left(\frac{B_s}{\bar{B}_n} - \frac{A_s}{\bar{A}_n}\right)\right\},\;
    \bar{A}_n = \frac{1}{n}\sum_{s=1}^{n}A_s,\;
    \bar{B}_n = \frac{1}{n}\sum_{s=1}^{n}B_s
\end{align}
We note that an additional symbol $\theta_{n,T}$ is introduced for possible finite population correction when relatively large number of elements are sampled without replacement (see below). Classical adjustment is to set $\theta_{n,T} = (1 - \frac{n-1}{T-1})$ \citep{Cochran1977}. For practical purposes when $T$ is large, often $\theta_{n,T}=1$ will suffice. In fact, knowledge of $T$ is not needed at all as long as it is large enough relative to $n$, which is useful because the total number of true clusters/blocks is not known in advance. Confidence interval can be computed based on the variance estimate of the above, which is
\begin{align} \label{eq:estimated_variance}
    \widehat{V}(\widehat{E})=\left( \frac{\bar{B}_n}{\bar{A}_n} \right)^2\frac{\theta_{n,T}}{n(n-1)}
    \sum_{s=1}^n \left( \frac{A_s^2}{\bar{A}_n^2} + \frac{B_s^2}{\bar{B}_n^2} - 2\frac{A_sB_s}{\bar{A}_n\bar{B}_n} \right).
\end{align}

The specific values for $A_s$, $B_s$, and $T$ in each of our representation is described in appendix \ref{sec:appendix}.

@@

The reason why \texttt{P\_cluster\_block} performs better than \texttt{P\_record} is due to the fact that \texttt{P\_record} has been derived in a generic way which applies to any performance metric which can be expressed in cluster form similar to \eqref{eq:lemma_1}. On the other hand, \texttt{P\_cluster\_block} relies on speific properties of precision. Among other things, this ensures that \texttt{P\_cluster\_block} is constrained to be between $0$ and $1$. In practice, \texttt{P\_cluster\_block} should be preferred as a pairwise precision estimator.

@@

\begin{table}\label{tab:precision_results}
    \centering
    \caption{\centering Bias and root mean squared error of precision estimators for the simulation study described in section \ref{sec:simulation_study}. The \texttt{rate} variable represents the percentage misattribution rate.}
\begin{tabular}{llccccccccc}
\toprule
{} & \hfill{\textbf{rate}} & \multicolumn{3}{c}{5} & \multicolumn{3}{c}{10} & \multicolumn{3}{c}{20} \\
\cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
{} & \hfill{\textbf{sample size}} & {100} & {200} & {400} & {100} & {200} & {400} & {100} & {200} & {400} \\
\midrule
\multirow[c]{4}{*}{\textbf{bias}} & P\_cluster\_block & -0.004 & -0.002 & -0.001 & -0.002 & -0.001 & 0.001 & -0.004 & -0.001 & -0.000 \\
 & P\_naive & 0.099 & 0.099 & 0.099 & 0.195 & 0.195 & 0.195 & 0.366 & 0.366 & 0.365 \\
 & P\_record & -0.002 & 0.013 & 0.009 & -0.001 & 0.011 & 0.008 & -0.001 & 0.009 & 0.006 \\
\midrule
\multirow[c]{4}{*}{\textbf{rmse}} & P\_cluster\_block & 0.045 & 0.033 & 0.021 & 0.041 & 0.038 & 0.026 & 0.063 & 0.052 & 0.034 \\
 & P\_naive & 0.099 & 0.099 & 0.099 & 0.195 & 0.195 & 0.195 & 0.366 & 0.366 & 0.365 \\
 & P\_record & 0.294 & 0.232 & 0.169 & 0.260 & 0.205 & 0.150 & 0.207 & 0.162 & 0.117 \\
\bottomrule
\end{tabular}
\end{table}

@@

Table \ref{tab:results} shows estimated pairwise precision and recall from benchmark datasets and from our two hand-curated datasets. Note that each estimate is associated with a given population of inventor mentions which is a subset of granted US patents since 1976. We focused on US patents granted since 1976 as this is the main data product of patentsview.org.

The choice of estimators for the results presented in table \ref{tab:results} is as follows.
Since our two hand-curated datasets (see section \ref{sec:hand-disambiguation}) were obtained by sampling inventor clusters with probabilities proportional to cluster sizes, we used the \texttt{P\_cluster\_block} and \texttt{R\_record} estimators with corresponding probability weights. 
For the Israeli benchmark dataset, we assumed that a single block of inventor clusters were sampled alltogether and used the corresponding estimators defined through \eqref{eq:lemma_3} with a single block sample. Note that no variance estimates can be given for single samples.
For \cite{Li2014}'s inventors benchmark, given that inventor clusters were originally obtained from a set of inventor curriculum vitae, we assumed that inventor clusters were sampled uniformly at random. As such, we used cluster block estimators with constant probability weights.

\begin{table}[!h]
    \label{tab:results}
    \centering
    \caption{\centering Estimated pairwise precision and recall (with estimated standard deviation) from our four benchmark datasets.}
    \begin{tabular}{lccc}
    \toprule
    \textbf{dataset} & \textbf{est. precision ($\hat \sigma$)} & \textbf{est. recall ($\hat \sigma$)} & \textbf{scope} \\
    \midrule
    Staff 1 & 86\% (5.6\%) & 94\% (2.1\%)  & 1976 -- Dec. 31, 2022 (US granted)\\
    Staff 2 & 88\% (5.6\%) & 95\% (2.0\%) & 1976 -- Dec. 31, 2022 (US granted)\\
    Israeli Benchmark & 79\% (NA) & 94\% (NA) & 1976 -- 1999 (US granted)\\
    \cite{Li2014}'s Benchmark & 91\% (2.7\%) & 91\% (5.0\%) & 1976 -- 2010 (US granted)\\
    \bottomrule
    \end{tabular}
\end{table}

@@

@@

Olivier Binette led the project and wrote the bulk of the manuscript. Youngsoo Baek provided bias adjustment and uncertainty quantification for ratio estimators. Aida York and Emma Hickerson carried out data collection by manually reviewing inventor clusters. Sarvo Madhavan was a technical advisor and contributed to code. Christina Jones was an advisor and project manager. All authors provided input on the manuscript.

@@

For this example, we considered the RLdata10000 dataset from \citep{}. This is a synthetic dataset containing 10,000 records with first name, last name, and date of birth attributes. There is noise in these attributes and a $10\%$ duplication rate. Ground truth identity is known for all records.

The disambiguation algorithm we consider matches records together if any of the following conditions is met:
\begin{itemize}
    \item records match on first name, last name, and birth year,
    \item records match on first name, birth day, and birth year, or
    \item records match on last name, birth day, and birth year.
\end{itemize}
Note that this is not at all a good disambiguation algorithm. It has $52\%$ precision and $83\%$ recall. However, it allows to to showcase the issue with non-adjusted precision computed on cluster samples.

In our experiment, we have repeated 5,000 times the following 3-steps process:
\begin{enumerate}
    \item First, 200 records were sampled and the ground truth clusters associated to them were recovered. This provided a ``benchmark" dataset which was used for evaluation.
    \item Second, a trivial precision estimate was obtained by computing precision over the benchmark dataset. That is, predicted cluster assignments were restricted to records which appear in the benchmark data and precision was compared for this. Most often than not, the result was an observation of $100\%$ precision.
    \item Third, we computed our proposed precision estimator which corresponds to lemma \ref{lemma:record_sampling_representation} and the estimator $\widehat P_{\text{block}}$ defined in equation \eqref{eq:P_block}, with blocks corresponding to clusters and with sampling probabilities $p_b \propto \lvert b \rvert$.
\end{enumerate}
The distributions of the these two precision estimates over the 5,000 repetitions are shown in figure \ref{fig:precision_problem}. Our proposed estimator is accurate and nearly unbiased, whereas the trivial precision estimates have almost nothing to do with actual algorithmic performance.

>>>>

\usepackage{xcolor}
\newcommand{\ob}[1]{{\color{purple}#1}}

@@

\author[1,2]{Olivier Binette}
\author[1]{Youngsoo Baek}
\author[2]{Sokhna A York}
\author[2]{Emma Hickerson}
\author[2]{Sarvo Madhavan}
\author[2]{Christina Jones}
\affil[1]{Department of Statistical Science, Duke University}
\affil[2]{American Institutes for Research}

@@

\begin{abstract}
This paper introduces a novel evaluation methodology for entity resolution algorithms. It is motivated by PatentsView.org, a \ob{U.S. Patents and Trademarks Office patent data exploration tool} which disambiguates patent inventors using an entity resolution algorithm. We provide a data collection methodology and tailored performance estimators which account for sampling biases. Our approach is simple, practical and principled -- key characteristics which allow us to paint the first representative picture of PatentsView's disambiguation performance. This is used to inform PatentsView's users of the reliability of the data and to allow the comparison of competing disambiguation algorithms.
\end{abstract}

\section{Introduction}
Entity resolution (also called record linkage, deduplication, or disambiguation) is the task of identifying records in a database which refer to the same entity. An entity may be a person, a company, an object or an event. Records are assumed to contain partially identifying information about these entities. When there is no unique identifier (such as a social security number) available for all records, entity resolution becomes a complex problem which requires sophisticated algorithmic solutions \citep{Herzog2007, Christen2012, Dong2015, Ilyas2019, Christophides2019, Christen2019, Papadakis2021, Binette2022a}.

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions to inventors, where a single person's name may be spelled in different ways on two applications, and where two different inventors with the same name may be difficult to distinguish. Much research has been done within the fields of economics, computer science and statistics to disambiguate inventor mentions \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

@@

The entity resolution evaluation problem is to extrapolate from observed performance on small samples to real performance in a database with millions of records. \cite{Wang2022} refers to this as bridging the ``reality-ideality gap" in entity resolution, where high performance on benchmark datasets often does not translate in the real world. Here, performance may be defined as any combination of commonly used evaluation metrics for entity resolution, such as precision and recall, cluster homogeneity and completeness, rand index, or generalized merge distance \citep{Maidasani2012}. These metrics can be computed on benchmark datasets for which we have a ground truth disambiguation. However, the key evaluation problem is to obtain estimates which are representative of performance on the full data for which no ground truth disambiguation is available. This is challenging for the following reasons.

First, entity resolution problems \ob{do not scale linearly}. While it may be easy to disambiguate a small dataset, the opportunity for errors grows \textit{quadratically} in the number of records. As such, we may observe good performance of an algorithm on a small benchmark dataset, while the true performance of on the entire data may be something else entirely.\footnote{This particular effect of dataset size in entity resolution is explored in \cite{Draisbach2013} in the context of choosing similarity thresholds.} This is a problem that PatentsView.org currently faces. Despite encouraging performance evaluation metrics on benchmark datasets, with nearly perfect precision and recall reported in the latest methodological report \citep{Monath2021}, the data science team at AIR observes lower real-world accuracy. \ob{This phenomenon is illustrated in example \ref{first_example} below.}

A second problem is large class imbalance in entity resolution \citep{Marchant2017}. \ob{Viewing entity resolution as a classification problem, the task is to classify record pairs as being a match or non-match. However,} among all pairs of records, only a small fraction (usually much less than a fraction of a percent) refer to the same entity. The vast majority of record pairs are not a match. This makes it difficult to evaluate performance through random sampling of record pairs.

@@

Our approach, detailed in sections \ref{sec:approach} and \ref{sec:proposed_estimators}, addresses these challenges by putting forward novel cluster-based expressions for performance metrics which reflect various sampling schemes. Each of these representations immediately suggests simple estimators which properly account for the above issues.

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we have carried out a simple experiment which is described in detail in appendix \ref{appendix:example}. In short, we considered evaluating a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark data sets are obtained and used. In this experiment, we known that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

@@

Pairwise precision and recall metrics were first reported in \cite{Newcombe1959}, with \cite{Bilenko2003, Christen2007} emphasizing the importance of precision-recall curves for algorithm evaluation. Now, there are issues with the use of pairwise precision and recall in entity resolution applications, such as the large relative importance of large clusters. As such, other clustering metrics have been proposed, including cluster precision and recall, cluster homogeneity and completeness, and generalized merge distances \citep{Michelson2009, Menestrina2010, Maidasani2012, Barnes2015}. Important practical issues regarding the use of aggregate metrics are discussed in \cite{Hand2018}.

\ob{While our work focuses on estimating pairwise precision and recall, the general approach also applies to any cluster based performance evaluation metrics, such as those described in \cite{Michelson2009}.}

@@

\ob{In contrast to the above, our approach to estimation is more practical than pairwise sampling and applies to any black-box disambiguation algorithms such as those used at PatentsView.}

@@

Our approach to estimating performance metrics is based on the use of benchmark datasets which are already existing or which can be collected in a cost-effective way. These datasets contain entity clusters corresponding to sampling records and recovering all associated instances, to directly sampling clusters, or to sampling blocks. For each of these sampling processes, we propose estimators which correct for the issues discussed in section \ref{sec:challenges}, are nearly unbiased, and easy to use in practice.

@@

\ob{Furthermore, our approach is novel. To our knowledge, we are the first to propose unbiased performance estimators based on cluster and block samples. Past work either ignored biases when computing precision and recall from benchmark datasets \citep{frisoli2018exploring, Monath2021, han2019disambiguating}, did not provide estimates for one of precision or recall \citep{mcveigh2019scaling}, or provided solutions that are tailored to very specific record linkage models \citep{Belin1995}. We provide the first general solution to entity resolution evaluation which does not rely on sampling record pairs and which applies to any disambiguation algorithm.}

@@

The rest of the paper is organized as follows. In section \ref{sec:data_methods}, \ob{we describe: motivating benchmark datasets, our data collection process for evaluation, the proposed estimators, and our simulation study used to validate the performance of our estimators.} Section \ref{sec:results} then shows the results: our performance estimates and results from the simulation study. Finally, section \ref{sec:discussion} summarizes the paper and discusses future research directions.

@@

In this section, \ob{we introduce: motivating datasets used at PatentsView, our hand-disambiguation methodology, our proposed performance metric estimators, and our simulation framework used to compare estimators.} \ob{Note that we focus on \textit{inventor} disambiguation throughout, rather than the related problems of assignee and location disambiguation.}

@@

\paragraph{Israeli Inventors Benchmark}
\cite{trajtenberg2008identification} disambiguated the U.S. patents of Israeli inventors which were granted between 1963 and 1999. A total of 6,023 Israeli inventors were identified for this time period with 15,310 associated patents.

@@

Based on an original dataset from \cite{gu2008march}, \cite{Li2014} disambiguated the patent history (between 1975 and 2010) of 95 U.S. inventors.

@@

In addition to the above benchmark datasets, \ob{we have carried out a data collection experiment. This was motivated by the evaluation of the current PatentsView inventor disambiguation using the estimators proposed in section \ref{sec:proposed_estimators}.}

In total, 100 inventors were sampled with probability proportional to their number of granted patents. This was done by sampling inventor mentions uniformly at random and recovering all patents for a given inventor. These inventor mentions were from U.S. patents granted between 1976 and December 31, 2021.

@@

Note that our data collection methodology is biased towards PantentsView's current disambiguation. \ob{Indeed, we do not expect the staff to have found all errors or all missing inventor mentions from predicted clusters. The staff used their best judgment, supported by a thorough search, to resolve inventor mentions.} In cases where no errors were found, the current disambiguation was assumed to be correct. Performance estimates based on this data might therefore be slightly optimistic. \ob{This should be acknowledged when reporting performance estimates to PatentsView.org users. Otherwise, for the purpose of improving the current disambiguation algorithm, the slightly biased data is still appropriate to use. It represents the most visible errors in the current disambiguation rather than the totality of them.}

@@

\begin{remark}
    Lemma \ref{lemma:record_sampling_representation} and lemma \ref{lemma:cluster_sampling_representation} can be generalized to apply to any performance metric which can be expressed as a sum $\sum_{c \in \mathcal{C}} h(c, \widehat{\mathcal{C}})$, for some function $h$, or as a function of such sums. For instance, the use of the so-called \textit{cluster} precision and \textit{cluster} recall \citep{Barnes2015}, or of cluster homogeneity and completeness \citep{Barnes2015}, can be more appropriate in the presence of large clusters. We leave these generalizations as extensions of our work.
\end{remark}

@@

All of the expressions for precision and recall in lemmas \ref{lemma:record_sampling_representation} -- \ref{lemma:block_sampling_representation} are either population means or ratios of population means. As such, they can be estimated using sample means and ratios of sample means. For readability, we present here a generic formula for an approximately unbiased estimator of the ratio of means, and then specify the needed quantities for each representation below. The estimator applies a first order bias correction to the ratio of sample means, based on a Taylor approximation. Approximate confidence intervals can be computed based on the variance estimator of the ratio of sample means by Taylor approximation, assuming the corrected estimator has small bias \citep{fuller2011sampling}. The generic formula for estimating the ratio of $T$-sized ``population'' (of records/clusters/blocks) means of the form
\begin{align}
    E=\frac{1/T\sum_{i=1}^{T} B_i}{1/T\sum_{i=1}^T A_i},
\end{align}
assuming we have sampled $n$ elements (records/clusters/blocks), is
\begin{align}\label{eq:bias_adjustment}
    \widehat{E} = \frac{\bar{B}_n}{\bar{A}_n} \left\{ 1 + \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \frac{A_s}{\bar{A}_n} \left(\frac{B_s}{\bar{B}_n} - \frac{A_s}{\bar{A}_n} \right) \right\}, \; \bar{A}_n = \frac{1}{n} \sum_{s=1}^{n} A_s, \; \bar{B}_n = \frac{1}{n} \sum_{s=1}^{n} B_s
\end{align}
We note that an additional symbol $\theta_{n,T}$ is introduced for possible finite population correction when relatively large number of elements are sampled without replacement (see below). Classical adjustment is set to $\theta_{n,T} = (1 - \frac{n-1}{T-1})$ \citep{Cochran1977}. For practical purposes when $T$ is large, often $\theta_{n,T}=1$ will suffice. In fact, knowledge of $T$ is not needed at all as long as it is large enough relative to $n$, which is useful because the total number of true clusters/blocks is not known in advance. Confidence interval can be computed based on the variance estimate of the above, which is
\begin{align} \label{eq:estimated_variance}
    \widehat{V}(\widehat{E}) = \left( \frac{\bar{B}_n}{\bar{A}_n} \right)^2 \frac{\theta_{n,T}}{n(n-1)} \sum_{s=1}^n \left( \frac{A_s}{\bar{A}_n} - \frac{B_s}{\bar{B}_n}\right)^2.
\end{align}
The specific values for $A_s$, $B_s$, and $T$ in each of our representation is described in appendix \ref{sec:appendix}.

@@

The reason why \texttt{P\_cluster\_block} performs better than \texttt{P\_record} is due to the fact that \texttt{P\_record} has been derived in a generic way which applies to any performance metric which can be expressed in cluster form similar to \eqref{eq:lemma_1}. On the other hand, \texttt{P\_cluster\_block} relies on specific properties of precision. Among other things, this ensures that \texttt{P\_cluster\_block} is constrained to be between $0$ and $1$. In practice, \texttt{P\_cluster\_block} should be preferred as a pairwise precision estimator.

@@

\begin{table}\label{tab:precision_results}
    \centering
    \caption{\centering Bias and root mean squared error (rmse) of precision estimators for the simulation study described in section \ref{sec:simulation_study}. The \texttt{rate} variable represents the percentage misattribution rate.}
\begin{tabular}{llccccccccc}
\toprule
{} & \hfill{\textbf{rate}} & \multicolumn{3}{c}{5} & \multicolumn{3}{c}{10} & \multicolumn{3}{c}{20} \\
\cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
{} & \hfill{\textbf{sample size}} & {100} & {200} & {400} & {100} & {200} & {400} & {100} & {200} & {400} \\
\midrule
\multirow[c]{4}{*}{\textbf{bias}} & P\_cluster\_block & -0.004 & -0.002 & -0.001 & -0.002 & -0.001 & 0.001 & -0.004 & -0.001 & -0.000 \\
 & P\_naive & 0.099 & 0.099 & 0.099 & 0.195 & 0.195 & 0.195 & 0.366 & 0.366 & 0.365 \\
 & P\_record & -0.002 & 0.013 & 0.009 & -0.001 & 0.011 & 0.008 & -0.001 & 0.009 & 0.006 \\
\midrule
\multirow[c]{4}{*}{\textbf{rmse}} & P\_cluster\_block & 0.045 & 0.033 & 0.021 & 0.041 & 0.038 & 0.026 & 0.063 & 0.052 & 0.034 \\
 & P\_naive & 0.099 & 0.099 & 0.099 & 0.195 & 0.195 & 0.195 & 0.366 & 0.366 & 0.365 \\
 & P\_record & 0.294 & 0.232 & 0.169 & 0.260 & 0.205 & 0.150 & 0.207 & 0.162 & 0.117 \\
\bottomrule
\end{tabular}
\end{table}

@@

Table \ref{tab:results} shows estimated pairwise precision and recall from benchmark datasets and from our two hand-curated datasets. Note that each estimate is associated with a given population of inventor mentions which is a subset of granted U.S. patents since 1976. We focused on U.S. patents granted since 1976 as this is the main data product of PatentsView.org.

The choice of estimators for the results presented in table \ref{tab:results} is as follows.
Since our two hand-curated datasets (see section \ref{sec:hand-disambiguation}) were obtained by sampling inventor clusters with probabilities proportional to cluster sizes, we used the \texttt{P\_cluster\_block} and \texttt{R\_record} estimators with corresponding probability weights. 
For the Israeli benchmark dataset, we assumed that a single block of inventor clusters was sampled and used the corresponding estimators defined through \eqref{eq:lemma_3} with a single block sample. Note that no variance estimates can be given for single samples.
For \cite{Li2014}'s inventors benchmark, given that inventor clusters were originally obtained from a set of inventor curriculum vitae, we assumed that inventor clusters were sampled uniformly at random. As such, we used cluster block estimators with constant probability weights.

\begin{table}[!h]
    \label{tab:results}
    \centering
    \caption{\centering Estimated pairwise precision and recall (with estimated standard deviation) from our four benchmark datasets.}
    \begin{tabular}{lccc}
    \toprule
    \textbf{dataset} & \textbf{est. precision ($\hat \sigma$)} & \textbf{est. recall ($\hat \sigma$)} & \textbf{scope} \\
    \midrule
    Staff 1 & 86\% (5.6\%) & 94\% (2.1\%)  & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Staff 2 & 88\% (5.6\%) & 95\% (2.0\%) & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Israeli Benchmark & 79\% (NA) & 94\% (NA) & 1976 -- 1999 (U.S. granted)\\
    \cite{Li2014}'s Benchmark & 91\% (2.7\%) & 91\% (5.0\%) & 1976 -- 2010 (U.S. granted)\\
    \bottomrule
    \end{tabular}
\end{table}

@@

\ob{One important topic for future work is the quantification of uncertainty associated with errors in the hand-disambiguation process. This is challenging problem given the lack of validation information available. Surveying inventors to validate the hand-disambiguation process would be one approach to explore for this. Sensitivity analyses involving potential errors in the hand-disambiguation process could also be informative. We refer the reader to \citep{bailey2017well} for a state-of-the-art hand-labeling study which evaluated human review accuracy.}

\ob{Another important topic for future work is the development of estimators for additional performance metrics. As we have seen in the simulation study, the generically-derived estimators (e.g., from lemma \ref{lemma:cluster_sampling_representation}) do not perform as well as estimators derived using specific properties of pairwise precision and recall (lemma \ref{lemma:block_sampling_representation}). As such, care should be taken to obtain efficient estimators for every metric of interest.}

@@

Olivier Binette led the project and wrote the bulk of the manuscript. Youngsoo Baek provided bias adjustment and uncertainty quantification for ratio estimators. Sokhna A York and Emma Hickerson carried out data collection by manually reviewing inventor clusters. Sarvo Madhavan was a technical advisor and contributed to code. Christina Jones was an advisor and project manager. All authors provided input on the manuscript.

@@

For this example, we considered the RLdata10000 dataset from \citep{RecordLinkage}. This is a synthetic dataset containing 10,000 records with first name, last name, and date of birth attributes. There is noise in these attributes and a $10\%$ duplication rate. Ground truth identity is known for all records.

The disambiguation algorithm we consider matches records together if any of the following conditions is met:
\begin{itemize}
    \item records agree on first name, last name, and birth year,
    \item records agree on first name, birth day, and birth year, or
    \item records agree on last name, birth day, and birth year.
\end{itemize}
Note that this is not at all a good disambiguation algorithm. It has $52\%$ precision and $83\%$ recall. However, it allows to to showcase the issue with non-adjusted precision computed on cluster samples.

In our experiment, we have repeated 5,000 times the following 3-steps process:
\begin{enumerate}
    \item First, 100 (resp. 200 and 400) records were sampled and the ground truth clusters associated to them were recovered. This provided a ``benchmark" dataset which was used for evaluation.
    \item Second, a trivial precision estimate was obtained by computing precision over the benchmark dataset. That is, predicted cluster assignments were restricted to records which appear in the benchmark data and precision was compared for this. Most often than not, the result was an observation of $100\%$ precision.
    \item Third, we computed our proposed precision estimator which corresponds to lemma \ref{lemma:record_sampling_representation} and the estimator $\widehat P_{\text{block}}$ defined in equation \eqref{eq:P_block}, with blocks corresponding to clusters and with sampling probabilities $p_b \propto \lvert b \rvert$.
\end{enumerate}
The distributions of the these two precision estimates over the 5,000 repetitions and for sample of sizes 100, 200, and 400 are shown in figure \ref{fig:precision_problem}. Our proposed estimator is accurate and nearly unbiased, whereas the trivial precision estimates have almost nothing to do with actual algorithmic performance.
